Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2022-08-04 supplement.tex
typo fix

In our preliminary models, the unconstrained inferred split times between East
African populations and Europeans was inferred to have occurred by around
$40$ka, which is more recent than we may have expected given the
well-documented presence of modern humans outside of Africa at that time
\citep[e.g.]{Hublin2020-nk,Hajdinjak2021-jo}, as well as previous
genetics-based dating of splits between African and Eurasian populations
\citep[e.g.]{Jouganous2017-pq,Kamm2020-vn}. The unconstrained inferred split
between East and West African populations occurred about $10$ka earlier. These
recent split dates seems to be at odds with archaeological evidence for earlier
Out-of-Africa (OOA) expansion, and with the estimated date for Neanderthal
admixture into OOA populations. The parameter uncertainty due to finite genomes
does not account for this difference: in the high-likelihood models, the $95\%$
confidence intervals for the East African-European split was roughly $38 \pm
5$ka, and for the East/West African split was $45 \pm 10$ka, with the
uncertainties reflecting 2 standard deviations according to bootstrap of
genomic regions.  Thus we expect these discrepancies to result from model
misspecification in the details of the divergence between the ancestors of OOA
and east African populations. Our models supposes a single split prior to the
OOA event followed by constant symmetric migrations between eastern and OOA
populations, which is certainly an oversimplification.

>>>>

In our preliminary models, the unconstrained inferred split times between East
African populations and Europeans was inferred to have occurred by around
$40$ka, which is more recent than we may have expected given the
well-documented presence of modern humans outside of Africa at that time
\citep[e.g.]{Hublin2020-nk,Hajdinjak2021-jo}, as well as previous
genetics-based dating of splits between African and Eurasian populations
\citep[e.g.]{Jouganous2017-pq,Kamm2020-vn}. The unconstrained inferred split
between East and West African populations occurred about $10$ka earlier. These
recent split dates seem to be at odds with archaeological evidence for earlier
Out-of-Africa (OOA) expansion, and with the estimated date for Neanderthal
admixture into OOA populations. The parameter uncertainty due to finite genomes
does not account for this difference: in the high-likelihood models, the $95\%$
confidence intervals for the East African-European split was roughly $38 \pm
5$ka, and for the East/West African split was $45 \pm 10$ka, with the
uncertainties reflecting 2 standard deviations according to bootstrap of
genomic regions.  Thus we expect these discrepancies to result from model
misspecification in the details of the divergence between the ancestors of OOA
and east African populations. Our models supposes a single split prior to the
OOA event followed by constant symmetric migrations between eastern and OOA
populations, which is certainly an oversimplification.
